---
title: "Chengdu_p2fa_basic"
author: "Katrina Li"
date: '2022-09-04'
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
chengdupath <- "/Users/kechun/GitHub/P2FA_Mandarin_py3/praat_followup/data_chengdu_p2fa"
chengdufiles <- list.files(chengdupath,".*dia.*",full.names = T)
# note: speficy na.strings to make sure all columns are read as numbers
# extract patterns: 
# - [a]? For Cantonese it can be 
chengdu <-tibble(File = chengdufiles) %>%
  tidyr::extract(File, c("paridsentence"),"(S[0-9]*dia[N]?[1]?[A-I]+[0-9][a]?)",remove = FALSE) %>%
  mutate(Data = lapply(File, read.delim, sep = "\t",na.strings = "--undefined--")) %>%
  unnest(Data) %>%
  select(-File) %>%
  mutate(parid = str_extract(paridsentence,"S[0-9]*"),
         sen_lab = str_extract(paridsentence,"[A-I]+[0-9][a]?")
         )

chengdu.clean <- chengdu %>%
  mutate(
    group = substr(sen_lab,1,nchar(sen_lab)-1),
    group = if_else(endsWith(group,"2"), substr(group,1,nchar(group)-1),group),
    group = as.factor(group),
    group_number = str_sub(sen_lab,-1),
    # group number has a suffix 'a', the group number is marked as '2a'
    group_number = if_else(group_number=="a", "2a",group_number),
    group_number = as.factor(group_number),
    case = case_when(character_lab == "小"~"Name(prefix)",
                     character_lab == "英"~"Name",
                     character_lab == "银"~"Name",
                     character_lab == "影"~"Name",
                     character_lab == "映"~"Name",
                     character_lab == "捏"~"Verb",
                     character_lab == "拿"~"Verb",
                     character_lab == "买"~"Verb",
                     character_lab == "卖"~"Verb",
                     character_lab == "狮"~"Object",
                     character_lab == "篮"~"Object",
                     character_lab == "椅"~"Object",
                     character_lab == "袜"~"Object",
                     character_lab == "子"~"Object(suffix)"),
    case = factor(case,levels = c("Name(prefix)", "Name","Verb","Object","Object(suffix)","Particle")),
    case_num = as.numeric(case),
    tone_type = case_when( group == "A" ~ "T1",
                           group == "B" ~ 'T2',
                           group == "C" ~ 'T3',
                           group == "D" ~ 'T4'),
    tone_type = as.factor(tone_type),
    interval = as.numeric(interval),
    f0mean = as.numeric(f0mean),
    f0min = as.numeric(f0min),
    f0max = as.numeric(f0max),
    f0min_point = as.numeric(f0min_point),
    f0max_point = as.numeric(f0max_point)
  ) 

chengdu.filter <- chengdu.clean %>%
  filter(group_number != 3 | interval < 10) %>% # exclude the phrase 'bushi...'
  filter(group_number != 3 | interval > 1) %>% # exclude 'shi'
  filter(paridsentence == "S2diaC4"| sen_lab == "B4" | group_number != 4 | interval > 7) %>% # exclude the context before 'lian'
  filter(paridsentence == "S2diaC4"| sen_lab == "B4" | group_number != 4 | interval < 18) %>% # exclude the context after 'lian' sentence
  filter(sen_lab != "B4" | interval > 14) %>% # exclude the context before 'lian' for B4
  filter(!is.na(case)) %>%
  filter(paridsentence != "S2diaC4" | interval != 13) # just one extra point

# Briefly check the number
chengdu.filter %>%
  group_by(parid, group, group_number) %>%
  summarise(count = n()) %>%
  filter(count != 15)

# not right point numbers (wrong way of filter):
#S2A3, A4, C4
#S28A4
#S4D4

chengdu.f0 <- chengdu.filter %>%
  pivot_longer(cols = starts_with("f0_"), names_to = "syltime", names_prefix = "f0_", values_to = "f0") %>%
  group_by(paridsentence,group, group_number) %>%
  mutate(sentime = row_number(),
         f0 = as.numeric(f0), #na value produced becaused of undefined value in makring
         f0_semi = f2st(f0,base=50),
         syltime = as.numeric(syltime)
         )

# Check when there is no f0 data
chengdu.f0 %>%
  filter(is.na(f0)) %>%
  select(parid, rhyme_lab, character_lab, sen_lab,syltime )
range(chengdu.f0$sentime)
```

## Plot figure
```{r}
plot_fun <- function(x,title){
  gg_colours <- c("#999999", "#E69F00", "#56B4E9","#009E73", "#F0E442", "#0072B2")
ggplot(x, aes(syltime, meanf0, group = group_number,color = group_number)) +
  geom_line(size = 1) +
  scale_color_manual(name = "Focus condition",
                       breaks = c("1","2","2a","3","4","5"),
                       labels = c("Broad focus","Wh-focus","Wh-focus-repeat","Subject focus","Even focus","Object focus"),
                       values = gg_colours) +
  labs(x="Normalized Time point", y = "Mean f0 in semitone") +
  ggtitle(title) +
  theme(axis.title = element_text(size = 10), 
        legend.title = element_text (size =10), 
        legend.text = element_text(size = 8),
        legend.position = "bottom",
        plot.title = element_text(size=10))+
    ylim(12,31) +
  facet_grid(group~case)
}

chengdu.f0.ave <- chengdu.f0 %>%
  group_by(sen_lab,group, group_number, case, case_num, syltime,sentime) %>%
  summarise(
    count = n(),
    meanf0 = mean(f0_semi, na.rm = TRUE),
    sdf0 = sd(f0_semi, na.rm = TRUE)
  )


plot_fun(subset(chengdu.f0.ave,group=="A"),"Chengdu Tone 1")
plot_fun(subset(chengdu.f0.ave,group=="B"),"Chengdu Tone 2")
plot_fun(subset(chengdu.f0.ave,group=="C"),"Chengdu Tone 3")
plot_fun(subset(chengdu.f0.ave,group=="D"),"Chengdu Tone 4")
```
T2 has very limited changes under focus. In addition, the downward tone shows clear downstep trend. - But we need to check if this is due to individual variation. If not, then this tone shows some interesting properties.

The other tones show changes. For T3 and T4, the pre-focus shows a clear 'straighting' tendency, that is, the contour being erased, and connecting the beginning and the end of the tonal contour.For T1, it's a straighting but not connecting the beginning and the end.

T1 and T3 show clear post-focus compression, but T4 less clear.

